### 2016-12-01 paper_v2_GEB.Rmd
Add files via upload

---
title: 'Spatial considerations in the analysis of biological conservation data: space is special'
author: "Alexis Comber^1^, Steve Carver^1^, Carol Ximena Garzon-Lopez^2^, Paul Harris^3^, Duccio Rocchini^4^"
output: 
  pdf_document: 
    keep_tex: yes
geometry: margin=3cm
---

^1^ School of Geography, University of Leeds, Leeds, LS2 9JT, UK

Email: {a.comber; s.j.carver} @leeds.ac.uk

^2^ Ecology and Dynamics of Human-influenced Systems Research Unit, University of Picardie Jules Verne, FR-80037 Amiens, France.

Email: c.x.garzon@gmail.com

^3^ Rothamsted Research, North Wyke, EX20 2SB, UK

Email: paul.harris@rothamsted.ac.uk

^4^ Research and Innovation Centre, Dpt. Biodiversity and Molecular Ecology, Fondazione Edmund Mach, 38010 S. Michele all'Adige, Italy

Email: duccio.rocchini@fmach.it

```{r functions, include=FALSE}
# A function for captioning and referencing figures
fig <- local({
    i <- 0
    ref <- list()
    list(
        cap=function(refName, text) {
            i <<- i + 1
            ref[[refName]] <<- i
            paste( text, sep="")
        },
        ref=function(refName) {
            ref[[refName]]
        })
})
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Normal Abstract
Geographically Weighted Regression (GWR) as first proposed by Brunsdon et al (1996) is a commonly used approach in many local geographical and biogeographical analyses. One of the of criticisms of the use of GWR is that local models may be subject to local collinearity between variables, even when none is observed globally. As yet none of the published research describing applications of GWR has addressed this issue. This paper does so. It applies GWR to the model suggested by Coudun et al (2006) to predict the presence of *Acer campestre* and to account for any spatial non-stationarity, as observed in many statistical patterns and relationships. The analysis tests for presence of local collinearity amongst predictor variables and applies a locally compensated GWR model where it is needed. The results describe the spatial variation of coefficient estimates predicting *Acer campestre* and compensates for any local collinearity amongst variables using a local ridge term. The paper discusses the need for *geography goggles* in  analyses of spatial and in so doing it encourages explicitly spatial ways of thinking. Such considerations are increasingly relevant as more and more data have a spatial component in the form of location, for instance from GPS, and all data are collected some*where*. These suggest the need for more informed approaches for analysing space and location than provided by standard statistical models.

# GEB Abstract
**Aim**: The aim of this research was to test the model suggested by Coudun et al (2006) that predicts the presence of *Acer campestre*, to explore species distributions models using Geographically Weighted Regression (GWR) as first proposed by Brunsdon et al (1996), and to take account of local collinearity amongst predictor variables. Local collinearity may occur even when none is observed globally. None of the many papers describing applications of GWR have addressed this issue.\
**Location**: UK \
**Time period**: 1980 and 2010\
**Major taxa studied**: *Acer Campestre*\
**Methods**: Data on the presence of *Acer Campestre* were downloaded from GBIF, and linked to data describing  autumn rainfall, wildness quality and potential evapotranspiration (PET). These data were aggregated over OS 10km squares. The analysis compares the results of predictive models from applying a standard OLS regression, a standard GWR and a locally compensated ridge GWR. A GWR diagnostics procedures was used to test for the presence of local collinearity and where found a locally compensated GWR model was applied.\
**Results**: The results show that the presence of *Acer campestre* is significantly associated with potential evapotranspiration and Wilderness Quality Index. The GWR analyses describe the spatial variation of coefficient estimates predicting *Acer campestre* and these are contrasted with the locally compensated GWR.\
**Main conclusions**: The paper discusses the need for *geography goggles* in  analyses of spatial and in so doing it encourages explicitly spatial ways of thinking. Such considerations are increasingly relevant as more and more data have a spatial component in the form of location, for instance from GPS, and all data are collected some*where*. These suggest the need for more informed approaches for analysing space and location than provided by standard statistical models.\

\newpage

# 1. Introduction

Ecological data increasingly have locational attributes thanks to near ubiquitous GPS in measurement and observation devices. This paper emphasises the importance of explicitly considering the spatial properties of data and argues that location cannot be treated as just another variable, precisely because of the prevalence of spatial autocorrelation or spatial non-stationarity in data, variables and statistical relationships. Space is special and requires *geography goggles* to be worn in applied biogeographical / ecological analyses and their associated methods and models. In order to demonstrate the value of spatially explicit methods, this paper tests the regression model suggested by Coudun et al (2006) but does this using a Geographically Weighted (GW) framework (Brunsdon et al, 1996), in this case a GW Regression (GWR). Coudun et al (2006) found the presence of *Acer campestre* in France to be significantly related to two factors: rainfall and evapotranspiration. Here Coudon’s analysis is extended to account for a particular form of spatial dependence in the data which occurs when changes in properties of nearby features are found to be correlated. This contradicts the underlying assumptions of independence and stationarity made in many statistical analyses and inferences. The result is spatial autocorrelation or spatial non-stationarity when the statistical pattern or statistical relationship observed in one location differs from that in another. To explore these, this paper applies a GWR analysis to examine the spatial variation in the relationships between the predictor variables and *Acer campestre* (field maple) presence to test for the presence of local, non-stationary relationships. GWR is directly concerned with modelling spatial non-stationarity  and in doing so will often indirectly account for any spatial autocorrelation (e.g. Harris etc al. 2011)

Spatially explicit approaches such as GWR reflect Tobler's First Law of Geography (Tobler, 1970). They seek to quantify *local* patterns in relationships and processes rather than global 'whole map' statistics. Typically,  local models are constructed from subsets of the data, for example by selecting data under a moving window or kernel, and local methods such as GWR have been used in a number of biodiversity and ecological studies. Recent examples include Roll et al (2015) who examined the spatial variation tree height as a predictor of species richness for different taxa, Wang et al (2016) who used to GWR to develop local models of net primary production in Chinese forests and Keith et al (2013) who sought to predict local species richness using a GWR analysis of environmental variables. However, local methods may be subject to collinearity even when it is not observed globally (Brunsdon et al 2012) because of the particular characteristics of the local subset. Collinearity will affect model reliability, result in inferential bias through unstable parameter estimates, inflated standard errors and difficulties in separating variable effects, etc (Dormann et al 2013; Meloun et al. 2002). Collinearity is commonly tested for in analyses of a-spatial data and may even be used to predict missing data. Despite the many research papers describing applications of GWR, as yet there are no instances of research that has considered the potential effects of local collinearity. As well as extending the analysis of Coudun et al (2006) using GWR, this paper explicitly tests for and compensates for locally collinearity where found, with dramatic differences between the coefficient estimates and the mapped GWR outputs without and without local compensations. The paper proceeds with a review of approaches for handling spatial auto-correlation (Section 2), Section 3 describes data and study area along with the issues of collinearity and local collinearity, before describing the results of a GWR analysis, a GW local collinearity diagnostics analysis and a locally compensated ridge regression in Section 4. Section 5 discusses the wider implications of the research int he context of the increasing availability of spatial data and some conclusions are drawn in Section 6.

# 2. Background

Understanding species distribution and spatial dependencies is a key concept in ecological research (Rocchini et al., 2015), in particular when dealing with biodiversity monitoring for conservation practices (Honrado et al., 2016). Several papers have dealt with the importance of explicitly taking space into account when trying to model the distribution of a certain species, and, overall, its change in space and time (see Pearman et al., 2008 and references therein). The description of the distribution of biodiversity at different spatial and temporal scales has long been the focus of ecology and biogeography. Reliable descriptions of species distributions are fundamental for conservation and research purposes (Dormann, 2007) and for conservation, where a lack of accuracy hamper the potential to provide effective tools for the development of spatially and temporally informed biodiversity management strategies (Guisan et al., 2014, Porfirio et al., 2014). However, when modelling species distributions, a number of statistical problems must be solved before spatial models can be built. These include the uncertainty related to the sampling of the modelled species (Rocchini et al., 2011), correlation among predictors (Dormann et al., 2012), potential variability over space of the predictors being used (Rocchini et al., 2016) and   spatio-temporal dependencies of the predictors (Zuur and Ieno, 2016). Among these, spatial non-stationarity and scale dependence have the greatest impact on the observed ecological processes and their relative patterns, such as the distribution of biodiversity over space (Foody, 2004).

The importance of spatial considerations has long been recognised in a number of disciplines: Fischer (1935) in crop science; Kolmogorov (1941), Gandin (1965) in meteorology; Krige (1951) and  Matheron (1963) in mining; Matérn (1960) in forestry; theoretical developments by Moran (1950) and Yaglom (1955); Berry and Marble (1968) and Chorley and Haggett (1967) in geography; and Legendre and Legendre (1991) in ecology.

The major issue with modelling spatial data is the lack of independence of the spatial objects and the assumption of stationarity in the processes being modelled. These are particularly true with socio-economic data, but are also true of bio-geographical, environmental and ecological data. Both concerns relate to Tobler's dictum (the first Law of Geography) that: '*everything is related to everything else, but near things are more related than distant things*' (Tobler 1970). However, in spite of numerous methodological advances in addressing such spatial modelling problems, policy related research is still routinely informed by non-spatial modelling practices, which are heroic at best, and ill-advised at worst. The lineage, from the late 1970s onwards, of such methodological advances can be loosely traced through spatial statistics texts from Journel and Huigbrechts (1978) to Cressie (1993), to Chilès and Delfiner (1999) and to Cressie and Wilke (2011).

Focusing on regression models, the main drawback in assuming observations are independent is that any spatial dependencies turn up in the residuals, as they are not explicitly added to the model. For area-based data, ways to account for this include Besag's (1974) conditionally autoregressive model or Anselin's (1988) spatially autoregressive model.  In both cases, a criterion is needed for determining 'nearby' and is often supplied by the entries in a matrix of adjacencies. Wall (2004) notes that spatial covariance structures are implied by such matrices, which sometimes act as confounders in the analysis implying counter-intuitive spatial processes. Vastly different covariance structures can be observed for different values of the autoregressive parameter. This is worrying for area-based modelling and is similar in nature to that found in Openshaw and Taylor's (1979) analysis of voting data in Iowa US, where different arrangements of the spatial units yielded correlations from -1 to +1 for the same pre-aggregated data. For point-based data, geostatistical regressions are commonly used, where spatial dependencies are modelled directly via the variogram, a function specifying a deterministic relationship between point pairs and their correlation. Recent advances have also seen geostatistical models applied in health studies where area-based data is common (Goovaerts 2009). These models are worthy in that they do not suffer from the pitfalls described by Wall (2004).

Spatial information can also be accounted for by expressing the regression's coefficients as functions of the spatial coordinates (Casetti 1982; Gorr and Oligschaeger 1994), permitting a model of relationship nonstationarity. A major advance in this area was advent of the GWR model (Brunsdon et al. 1996). Here spatial structures in the data are incorporated into the model through a continuous distance-decay weighting scheme. The spatial extent of the scheme is controlled through the bandwidth parameter, which is optimised to give the best model fit to the data. The resultant localised regression coefficients are mapped to display their spatial variation.  Brunsdon et al. (1998) and  Harris et al. (2015) provide some simple inferential mechanisms to determine whether the coefficients exhibit significant spatial variation; and other notable advances in the GWR method can be found in Nakaya et al. (2005), Wheeler (2007), Huang et al. (2010), Harris et al. (2011), Brunsdon et al. (2012) and Silva and Fotheringham (2015). However, a fully-coherent inferential structure of GWR remains to be developed. It may be that, given GWR's kernel smoothing roots (e.g. see Cleveland 1979), the technique is best reserved for use as an exploratory and highly visual tool only, to act as a valuable precursor to a more sophisticated spatially-varying coefficient model, such as those proposed by Gelfand et al. (2003) or Assunção (2003).

The over-riding issue is that dependence and some form of nonstationarity are endemic in spatial data, and model forms should be made widely available which allow users to explore these structures both pre- and post-modelling. When spatial non-stationarity holds, the predictor and response variables is expected to change over space in an area.  Under spatial non-stationarity the relationship between the distribution of a certain species and the predictors shaping it may change over space. Thus any modelling activity  should also incorporate mechanisms to handle these characteristics of spatial data. This is an important challenge and this paper focuses on the modelling of spatial nonstationarity via the GWR model in order to demonstrate how to address process heterogeneity in relationships. The GWR model itself is formally described in Section 3, both in a basic form and an adapted form to deal with potential problems of localised collinearity in the predictor variables (Brunsdon et al. 2012; Gollini et al. 2015).

# 3. Methods

```{r eval=T, echo=F, message=F, warning=F, cache = TRUE}
# you will only need to run the lines below after 
# you have installed the packages on your computer
library(GISTools)
library(rgdal)
library(GWmodel)
library(repmis)
library(knitr)

# Part 1. load and prepare the data

# 1.1 load prepared data files

#### **** You can download the data from the Github site *****
#### **** And then run source_data as below *****
# source_data("https://github.com/lexcomber/SpEcolCons2016/blob/master/SpEcolCons.RData?raw=True")

#### ***** OR load the data from a local file *****
load('~/Desktop/my_docs_mac/leeds_work/research/conservation/anal/SpEcolCons.RData')

# Acknowldgements Rainfall
# https://catalogue.ceh.ac.uk/documents/f2856ee8-da6e-4b67-bedb-590520c77b3c
# You must cite: Tanguy, M.; Dixon, H.; Prosdocimi, I.; Morris, D. G.; Keller, V. D. J. (2015). Gridded estimates of daily and monthly areal rainfall for the United Kingdom (1890-2014) [CEH-GEAR]. NERC Environmental Information Data Centre. http://doi.org/10.5285/f2856ee8-da6e-4b67-bedb-590520c77b3c

# Acknowldgements PET
#http://csi.cgiar.org/Aridity/
#Trabucco, A., and Zomer, R.J. 2009. Global Aridity Index (Global-Aridity) and Global Potential Evapo-Transpiration (Global-PET) Geospatial Database. CGIAR Consortium for Spatial Information. Published online, available from the CGIAR-CSI GeoPortal at: http://www.csi.cgiar.org.
# clip PET to OS

# Acknowledgement Wildness Quality Index
# Kuiters, A. T., van Eupen, M., Carver, S., Fisher, M., Kun, Z., & Vancura, V. (2013). Wilderness register and indicator for Europe. Final Report - http://ec.europa.eu/environment/nature/natura2000/wilderness/pdf/Wilderness_register_indicator.pdf

# 1.2 interpolate the over OS 10k grid
# wqi at 1k 
# pet at 30 arc second (1k at equator) 
# rain at 1k 
# everything projected to OS

# Aggregate the single variable layers
wqi.agg <- aggregate(wqi[,1], os10, mean)
pet.agg <- aggregate(pet[,1], os10, mean)
rain.agg <- aggregate(rain[,1], os10, mean)

# convert to points
wqi.agg <- SpatialPointsDataFrame(coordinates(wqi.agg), 
	data = wqi.agg@data, proj4string = os.proj)
pet.agg <- SpatialPointsDataFrame(coordinates(pet.agg), 
	data = pet.agg@data, proj4string = os.proj)
rain.agg <- SpatialPointsDataFrame(coordinates(rain.agg), 
	data = rain.agg@data, proj4string = os.proj)

# Aggregate the rainfall data from 'data'
# define a summary function
year.func <- function(x){
	sum(x, na.rm=T)/mean(x, na.rm = T)
}
years <- sort(unique(data.sp$year))
res.vec <- matrix(data = 0, ncol = length(years), nrow = nrow(os10))
for (i in 1:length(years)){
	year.i <- years[i]
	index.i <- which(data.sp$year == year.i)
	data.i <- data.sp[index.i,]
	data.i <- spTransform(data.i, os.proj)
	agg.i <- aggregate(data.i["year"], os10, year.func)
	res.vec[, i] <- agg.i$year
}
colnames(res.vec) <- years
index <- is.na(res.vec)
res.vec[index] <- 0
# create the aggregated variable
tree.agg <- SpatialPointsDataFrame(os10, 
	data.frame(tree = rowSums(res.vec)), 
	proj4string = os.proj)

# Uncomment to check dimensions
# dim(wqi.agg)
# dim(rain.agg)
# dim(pet.agg)
# dim(tree.agg)

# 1.3 create a single dataset for use in the analyses
df <- data.frame(tree = tree.agg@data, pet = pet.agg@data, 
	rain = rain.agg@data, wqi = wqi.agg@data)

d.a <- SpatialPointsDataFrame(coordinates(os10), data = data.frame(df))
# summary(d.a)
# get rind of any NAs
index<- !is.na(rowSums(d.a@data))
d.a <- d.a[index,]
```

## 3.1 Data and Study Area

Data recorded between 1980 and 2010 describing the presence of *Acer campestre* was downloaded from GBIF using the `dismo` R package, and subsetted for the UK. The data contained 22,701 records whose spatial distribution are shown in Figure 1, with a median of `r summary(colSums(res.vec))[3]`  records per year, a 1st quartile of `r summary(colSums(res.vec))[2]`  and a `r summary(colSums(res.vec))[4]` quartile of 917 records. The data for all years were summed over Ordnance Survey 10km grids because of the uneven distribution in time and space. Potential alternative approaches including generating absence points, background data (Phillips et al. 2009) to charcaterise study area environments or pseudo-absences (e.g. VanDerWal et al., 2009), indicating where absences might occur. However, pseudo absence approaches require a number of assumptions and lack statistical methods for handling the overlap between presence and background points (Ward et al. 2009; Phillips and Elith, 2011), absence data may be biased and / or incomplete (Kery et al., 2010) and background data approaches generate the same measures irrespective of where the species is observed (Hijmans and Elith, 2015).

```{r, fig.width=9, echo = F, eval=T,  cache = T, message=F, warning=F, fig.cap=fig$cap("", "a) The raw data points with a transparency term to show density of points, and b) Data summed over OS 10km grid cells.")}
par(mfrow = c(1,2))
par(mar = c(0,0,1,0))
plot(os10m, lwd = 0.5)
plot(data.sp[os10m,], pch = 1, cex = 0.2, col = "#2525254C", add = T)
title("a)")
sh <-auto.shading(d.a$tree[d.a$tree>0], n = 7, cols=brewer.pal(7,'Reds'))
sh$breaks <- c(1,5,10,20,30,50)
choropleth(d.a, v = d.a$tree, sh, pch = 19, cex = 0.4, add = F)
plot(os10m, lwd = 0.5, add = T)
choro.legend(px = "topright", sh = sh, title("Trees per 10km sq"), cex = 0.7)
title("b)")
```

The study by Coudun et al (2006) found the presence of *Acer campestre* to be significantly related to Autumn rainfall and actual Thornthwaite evapotranspiration. In this study rainfall, Potential evapotranspiration (PET) and data describing a wildness quality index were used to construct a series of models for predicting the density of *Acer campestre* occurrence. Data on rainfall were downloaded from the NERC Environmental Information Data Centre (Tanguy et al, 2015) which provides monthly 1km estimated rainfall data for each year. The average Autumn (3 month) rainfall was calculated for each 1km. Mean annual PET data were downloaded from the CGIAR Consortium for Spatial Information (Trabucco and Zomer, 2009) which provides global data at 0.0083 degrees, approximately 1km. Finally, a wilderness quality dataset was included in the model. This was to explore the degree to the presence of *Acer campestre* may be related to anthropogenic disturbance – anecdotally this species is frequently used as an ornamental tree and found in field margins and hedges. The Wilderness Quality Index (WQI) data were generated for the whole of Europe at 1km resolution as described in Kuiters et al (2013). WQI can be considered as measure of non-anthropogenic activity and is based on a weighted linear combination of attributes related to wild land, wilderness and wildness (Carver et al, 2012, Comber et al, 2010) including naturalness of vegetation patterns, remoteness from settlements and human influence, and remoteness from roads. Each of these datasets were spatially aggregated over the OS 10m grid cells to generate mean values as shown in Figure 2.

```{r, fig.width=9, echo = F, eval=T,  cache = T, message=F, warning=F, fig.cap=fig$cap("", "The data used to construct the species model, a) mean monthly Autumn Rain (mm), b) mean annual potential evapotranspiration (PET), and c)  Mean Wildness Quality Index.")}
par(mfrow = c(1,3))
par(mar = c(0,0,1,0))

plot(os10m, lwd = 0.5)
sh <-auto.shading(d.a$rain[d.a$rain > 0 ], n = 7,
	cols=brewer.pal(7,'Blues'))
choropleth(d.a, v = d.a$rain, sh, pch = 19, cex = 0.3, add = T)
choro.legend(px = "topright", sh = sh, cex = 1)
title("a)", cex.main = 1.5)

plot(os10m, lwd = 0.5)
sh <-auto.shading(d.a$pet[!is.na(d.a$pet)], n = 7,
	cols=brewer.pal(7,'YlOrRd'))
choropleth(d.a, v = d.a$pet, sh, pch = 19, cex = 0.3, add = T)
choro.legend(px = "topright", sh = sh, cex = 1)
title("b)", cex.main = 1.5)

plot(os10m, lwd = 0.5)
sh <-auto.shading(d.a$wqi[!is.na(d.a$wqi)], n = 7,
	cols=brewer.pal(7,'GnBu'))
choropleth(d.a, v = d.a$wqi, sh, pch = 19, cex = 0.3, add = T)
choro.legend(px = "topright", sh = sh, cex = 1)
title("c)", cex.main = 1.5)
```

## 3.2 Analysis

A multi-stage analysis was applied to model *Acer campestre* distributions. First, exploratory analyses were undertaken using a standard OLS regression to model distributions as an initial step. This identified significant predictor variables, under the assumption that relationships between predictor variables (rainfall, PET and wilderness) and species distributions are stationary (i.e. global). Then a GWR analysis was applied to examine the spatial variation in the relationships between the predictor variables and *Acer campestre* distributions (i.e. to test for the presence of local, non-stationary relationships). In overview, GW approaches use a moving window or kernel that passes through the study area. At each location being considered, data under the window are used to make a local calculation of some kind, such as a regression. The data are weighted by their distance to the kernel centre and in this way GW approaches construct a series of models at discrete locations in the study area. This is in contrast to global models, that consider all of the data (usually) in a single analysis of all data in the study area.

Next, the presence of global and local collinearity amongst the predictor variables was tested. This is a critical step in any regression, but for a GWR analysis it is usually overlooked. Collinearity occurs when variables exhibit linear or near linear relationships. Strong collinearity will affect model reliability and precision, generate unstable parameter estimates, inflated standard errors and inferential biases (Dormann et al 2013), and there may be problems in separating variable effects (Meloun et al. 2002). In a GWR analyses, collinearity may occur locally, with the construction of localised regressions, even when it is not observed globally (Wheeler and Tiefelsdorf, 2005; Wheeler 2007, 2009, 2013; Brunsdon et al 2012). A number of approaches exist to address collinearity in regression modelling, such as partial least squares regression, principal component analysis regression and ridge regression (Hoerl 1962; Hoerl and Kennard 1970). Ridge regression is a penalised model where extensions, such as the lasso and the elastic net also provide predictor variable sub-set selection (e.g. Zou and Hastie 2005). All such models could be adapted to a localised form and Wheeler (2007; 2009) has proposed both ridge and lasso versions of GWR to address any detrimental local collinearity effects. A related, locally-compensated ridge GWR model (LCR-GWR) is detailed in Brunsdon et al (2012), Lu et al. (2014) and Gollini et al (2015). It has advantages over the ridge GWR model of Wheeler (2007) in that a ridge term is applied locally and not globally. These studies also describe associated local collinearity diagnostics for GWR, such as the use of local correlations amongst pairs of predictors, local Variance Inflation Factors (VIFs) for each predictor, local variance decomposition proportions (VDPs) and the local condition numbers (CNs). The key point about LCR-GWR, is that a local ridge term is only applied where it is needed – when the local CN is above a pre-specified value in this case 30, which is a standard heuristic.

This study undertook a GWR analysis, with a locally-compensated ridge term if necessary, over a 200m grid of points covering the study area, with the aim of examining the spatial distribution of coefficient estimates predicting *Acer campestre* distributions and their spatial variation. Euclidean distances were used to weight data points under the kernel. These distances better reflect the spatial processes and relationships in environmental systems than network distance (Comber et al, 2008). For the kernel, an adaptive bi-square weighting function was applied, although a number of kernel functions can be specified for GW models as discussed in Gollini et al (2015). This generates higher weights at locations very near to the kernel centre relative to those towards the edge. For each data point ($P_j$) under the kernel (with a given bandwidth), a weight $w_{i,j}$ is calculated based on its distance to the centre of the kernel ($K_{i}$) as follows:

\begin{center}
\begin{equation}
w_{i,j} = 1 - ((d_{i,j})^2 / b^2)
\end{equation}
\end{center}
where $d_{i,j}$ is the distance in metres from the centre of the kernel $K_i$ to the data point $P_j$ and $b$ is the bandwidth.

An optimum kernel bandwidth for GWR can be found by minimising a model fit diagnostic. Options include a leave-one-out cross-validation (CV) score (Bowman 1984; Brunsdon et al. 1996). This optimises model prediction accuracy and the Akaike Information Criterion (AIC) (Akaike 1973; Fotheringham et al. 2002) optimises model parsimony by trading off prediction accuracy and complexity. In this case, the CV approach was applied to specify all GWR models, all using the bi-square weighting kernel and distances between locations.

The standard GWR model is:

\begin{center}
\begin{equation}
y_{i} = \beta_{i0} + \sum_{m}^{k=1}\beta_{ik}x_{ik} + \epsilon_i 
\end{equation}
\end{center}

where $y_i$ is the response variable at location $i$, $x_{ik}$ is the value of the $k^{th}$ predictor variable at location $i$, $m$ is the number of predictor variables, $\beta_{i0}$ is the intercept term at location $i$, $\beta_{ik}$ is the local regression coefficient for the $k^{th}$ predictor variable at location $i$ and $\epsilon_i$ is the random error at location $i$. The result of the weighting means that data nearer to the kernel centre make a greater contribution to the estimation of local regression coefficients at each local regression calibration point $i$.

## 3.3 Code

All of the analyses and mappings were undertaken in R, the free open source statistical software. The RMarkdown script used to produce this manuscript, including all the code used in the analysis and to produce the mapped figures, can be found at [https://github.com/lexcomber/SpatEcolPap](https://github.com/lexcomber/SpatEcolPap)

# 4. Results

## 4.1 Exploratory Regressions

A standard OLS regression models was undertaken and the resultant coefficient estimates and significance values are shown in Table 1 below. PET and Wilderness were found to be significant predictors of *Acer campestre* distributions at the 5% level (i.e. with a less than 95% chance of occurring randomly). Interestingly, in contrast to the findings of Coudun et al (2006), mean Autumn rainfall was not found to be significantly associated with the *Acer campestre* distributions.

```{r eval=T, echo=F, message=F, warning=F, cache = TRUE}
m <- lm(tree~pet+rain+wqi, data = data.frame(d.a))
#summary(m)
#coef(m)
tab <- summary(m)[4]
tab <- tab[[1]]
tab <- round(tab, 3)
rownames(tab) <- c("Intercept", "Mean annual PET", "Mean Autumn rainfall","Mean Wilderness Quality Index")
```

\begin{center}
{Table 1. The global regression co-efficient estimates.}
\end{center}

```{r eval=T, fig.width=5, echo=F, message=F, warning=F}
library(knitr)
kable(tab, booktabs = T, format = 'markdown') 
```

```{r eval=T, cache = T, results = 'hide', echo=F, message=F, warning=F}
bw2 <- bw.gwr(d.a$tree~d.a$pet+d.a$rain+d.a$wqi, data = d.a, kernel = "bisquare", adaptive = TRUE, approach = "CV") 
# create GW regression points
hg <- spsample(os10m,3500,'hexagonal',offset=c(0.5,0.5)) # 7500 later?

# create gwr.mod 
m.g2 <- gwr.basic(d.a$tree~d.a$pet+d.a$rain+d.a$wqi, data = d.a, regression.points = hg, bw = bw2, kernel = "bisquare", adaptive = TRUE)

# have a look at the model summary
#m.g2
```

```{r eval=T, echo=F, message=F, warning=F}
tab2 <- (m.g2$SDF@data[,1:4])
tab2 <- round(apply(tab2, 2, fivenum), 3)
rownames(tab2) <- c("Min", "1st Qu", "Median", "3rd Qu", "Max")
tab2 <- t(tab2)
rownames(tab2) <- c("Intercept", "Mean annual PET", "Mean Autumn rainfall","Mean Wilderness Quality Index")
```

\begin{center}
{Table 2. The variation of the coefficient estimates arising from a GWR analysis.}
\end{center}

```{r eval=T, echo=F, message=F, warning=F}
library(knitr)
kable(tab2, booktabs = T, format = 'markdown') 
```

The underlying theoretical framework provided by GWR tests for non-stationarity in processes and relationships between factors. A standard GWR analysis was undertaken and in this case an optimal adaptive bandwidth of `r bw2` data points was determined using a cross-validation procedure.

The local coefficient estimates from this GWR model are shown in Table 2 and the significant variables, PET and WQI are mapped in Figures 3 and 4, respectively. They indicate considerable variation around the median in the degree to which increases in the predictor variables are associated with *Acer campestre* distributions. For example, considering the inter-quartile ranges shows that, in some places:

- An increase in PET of 100 values is associated with a decrease of `r tab2[2,2] * 100` trees; 
- That each increase of 0.3 in the wilderness index is associated with a decrease of 1 tree (`r tab2[4,2]` * 0.3); 
But in other locations: 
- A decrease in PET of 100 values is associated with an increase of `r tab2[2,4] * 100` trees; 
-	That each increase of 0.5 in the wilderness index is associated with an increase of 1 tree (`r tab2[4,4]` * 0.5).

The local variation in coefficient estimates in Table 2 is in contrast to the global coefficient estimates in Table 1.

```{r eval=T, cache = T, echo=F, message=F, warning=F}
# convert the gw.model SDF to a Spatial Points Data Frame
library(GISTools)
m.gr2 = SpatialPointsDataFrame(m.g2$SDF,data.frame(m.g2$SDF), proj4string = os.proj)
```

```{r, fig.width=9, echo = F, cache = T, eval=T,  message=F, warning=F, fig.cap=fig$cap("", "The spatial distribution of the mean annual PET coefficient estimates, with the context of the original PET and species data.")}
# PET and GW coefs for PET

par(mar = c(1,1,1,0))
par(mfrow = c(1,3)) 
cx <- 0.4
sh <-auto.shading(d.a$pet[!is.na(d.a$pet)], n = 6,
	cols=brewer.pal(7,'YlOrRd')[2:7])
plot(os10m, lwd = 0.5)
choropleth(d.a, v = d.a$pet, sh, pch = 19, cex = cx, add = T)
choro.legend(px = "topleft", sh = sh, box.col = NA, cex = 0.7)
title("Mean annual PET")

plot(os10m, lwd = 0.5)
plot(data.sp[os10m,], add = T, cex = 0.4, pch = 1, col = "#2525254C")
title("Data Points")

sh = auto.shading(m.gr2$d.a.pet, n = 7, 
	cols = brewer.pal(7, "Spectral"))
plot(os10m, lwd = 0.5, add = F)
choropleth(m.gr2, m.gr2$d.a.pet, sh, pch = 19, cex = cx, add = T)
txt=expression(paste("GW Coeff: ",B[PET]))
choro.legend(px = "topright", sh = sh, box.col = NA, fmt = "%2.2f", cex = 0.7, title = txt)
tit <- sprintf("PET Global Coeff: %s",round(coef(m)[2], 3))
title(tit)
plot(os10m, lwd = 0.5, add = T)
```

```{r, fig.width=9, echo = F, cache = T, eval=T,  message=F, warning=F, fig.cap=fig$cap("", "The spatial distribution of the Wilderness Quality Index  coefficient estimates, with the context of the original WQI and species data.")}

# WQI and GW coefs for WQI
par(mar = c(1,1,1,1))
par(mfrow = c(1,3)) 
cx <- 0.4
sh <-auto.shading(d.a$wqi[!is.na(d.a$wqi)], n = 6,
	cols=brewer.pal(7,'YlOrRd')[2:7])
plot(os10m, lwd = 0.5)
choropleth(d.a, v = d.a$wqi, sh, pch = 19, cex = cx, add = T)
choro.legend(px = "topleft", sh = sh, box.col = NA, cex = 0.7)
title("Wildness Quality Index")

plot(os10m, lwd = 0.5)
plot(data.sp[os10m,], add = T, cex = 0.4, pch = 1, col = "#2525254C")
title("Data Points")

sh = auto.shading(m.gr2$d.a.wqi, n = 7, 
	cols = brewer.pal(7, "Spectral"))
plot(os10m, lwd = 0.5, add = F)
choropleth(m.gr2, m.gr2$d.a.wqi, sh, pch = 19, cex = cx, add = T)
txt=expression(paste("GW WQI: ", B[PET]))
choro.legend(px = "topright", sh = sh, box.col = NA, fmt = "%2.2f", cex = 0.7, title = txt)
tit <- sprintf("WQI Global Coeff: %s",round(coef(m)[4], 3))
title(tit)
plot(os10m, lwd = 0.5, add = T)
```

## 4.2 GWR Local Collinearity Diagnostics

The potential for detrimental effects due to local collinearity has been ignored by nearly all of the GWR analyses reported in the literature, regardless of domain or subject. Collinearity occurs when one predictor variable has a strong positive or negative relationship with another, typically when it is less than -0.8 or greater than +0.8. Critically, collinearity may be absent when calculated globally (i.e. from all the data values), but may be present locally when a subset of the data is considered, as is the case with a GWR analysis. Table 3 shows the results of evaluating collinearity globally and locally using the GWR collinearity diagnostics tool included in the `GWmodel` R package.

Globally, the Variance Inflation Factors (VIFs) are all less than 10, although 2 of the Variance Decomposition Proportions (VDPs) are greater than 0.5 and the Condition Number (CN) is greater than 30, using standard heuristics from Belsley et al (1980) and O’Brien (2007). Together these suggest the presence of variable collinearity when evaluated globally, considering all data points together. Applying GWR collinearity diagnostics to the GWR model above generates local VIFs, local VDPs and local CNs at the same scale (i.e. using the same adaptive bandwidth of 21 data points). The results in Table 3 indicate a high degree of *local* collinearity in the GWR model. These values suggest that the application of a LCR-GWR is warranted. The local collinearity measures are mapped in Figure 5.

```{r eval=T, echo=F, message=F, warning=F}
library(GWmodel)
library(car)
```

```{r eval=T, cache = T, echo=F, message=F, warning=F}
# Examine LOCAL Collinearity
gw.col <- gwr.collin.diagno(d.a$tree~d.a$pet+d.a$rain+d.a$wqi, data = d.a, bw = bw2, kernel = "bisquare", adaptive = TRUE)
```

```{r eval=T, cache = T, echo=F, message=F, warning=F}
gw.col.glob <- gwr.collin.diagno(d.a$tree~d.a$pet+d.a$rain+d.a$wqi, data = d.a, bw = nrow(d.a), kernel = "boxcar", adaptive = TRUE)
```

```{r eval=T, echo=F, message=F, warning=F}
### VIFs greater than 10 -> suggestive of collinearity
### CNs greater than 30 -> suggestive of collinearity
# summary(gw.col$SDF)
# VIFs & CNs for the variables considered
# summary(gw.col$SDF[,1:4])
tab3 <- gw.col$SDF@data[,c(1:4, 6:8)]
tab3 <- round(apply(tab3, 2, fivenum), 3)
rownames(tab3) <- c("Local Min", "Local 1st Qu", "Local Median", "Local 3rd Qu", "Local Max")
tab3 <- t(tab3)
row.i <- gw.col.glob$SDF@data[,c(1:4, 6:8)]
row.i <- round(apply(row.i, 2, fivenum), 3)[1,]
tab3 <- cbind(row.i, tab3)
colnames(tab3)[1] <- "Global"
rownames(tab3) <- c("VIF PET", "VIF Rainfall","VIF WQI", "CN", "VDP PET", "VDP Rainfall", "VDP WQI")
tab3 <- tab3[c(4,1:3,5:7),]
```

\begin{center}
{Table 3. Global and local collinearity measures: Condition Numbers with Variance Inflation Factors (VIFs) and Variance Decomposition Proportions (VDPs) for each predictor variable.}
\end{center}

```{r eval=T, echo=F, message=F, warning=F}
library(knitr)
kable(tab3, booktabs = T, format = 'markdown') 
```

```{r, fig.width=9, echo = F, eval=T, cache = T,  message=F, warning=F, fig.cap=fig$cap("", "The spatial distribution of the local CNs and the mean annual PET VIF and the mean Wilderness Quality Index VIF.")}

par(mar = c(0,1,1,1))
par(mfrow = c(1,3)) 
cx <- 0.5

sh = auto.shading(gw.col$SDF@data[,4], n = 7, 
	cols = brewer.pal(7, "Reds"))
plot(os10m, lwd = 0.5, add = F)
choropleth(gw.col$SDF, v = gw.col$SDF@data[,4], sh, pch = 19, cex = cx, add = T)
title("Local CN: all > 30!")
choro.legend(px = "topright", sh = sh, box.col = NA, fmt = "%2.0f", cex = 0.7, title = "CN")
plot(os10m, lwd = 0.5, add = T)

sh = auto.shading(gw.col$SDF@data[,1], n = 5, 
	cols = brewer.pal(5, "Blues"))
sh$breaks <- c(2,5,7,10)
plot(os10m, lwd = 0.5, add = F)
choropleth(gw.col$SDF, v = gw.col$SDF@data[,1], sh, pch = 19, cex = cx, add = T)
title("Local VIF (PET): some > 10")
plot(os10m, lwd = 0.5, add = T)
choro.legend(px = "topright", sh = sh, box.col = NA, fmt = "%2.1f", cex = 0.7, title = "VIF")

sh = auto.shading(gw.col$SDF@data[,3], n = 5, 
	cols = brewer.pal(5, "Greens"))
sh$breaks <- c(2,5,7,10)
plot(os10m, lwd = 0.5, add = F)
choropleth(gw.col$SDF, v = gw.col$SDF@data[,3], sh, pch = 19, cex = cx, add = T)
title("Local VIF (WQI): some > 10")
plot(os10m, lwd = 0.5, add = T)
choro.legend(px = "topright", sh = sh, box.col = NA, fmt = "%2.1f", cex = 0.7, title = "VIF")
```

## 4.3 Final GWR analysis

```{r eval=T, cache = T, echo=F, results = 'hide', message=F, warning=F}
# Compensate for local collinearity
# using GW locally comnsated ridge regression function
# First determine the new bandwidth
# This tries to find the bandwidth to fit a ridge term in every location where the CN was greater than 30 
lcr.bw <- bw.gwr.lcr(d.a$tree~d.a$pet+d.a$rain+d.a$wqi, data = d.a, kernel = "bisquare", adaptive = TRUE, lambda.adjust = TRUE, cn.thresh = 30)
```

```{r eval=T, cache = T, echo=F, message=F, warning=F}
# Then run the locally compensated GWR 
lcr.m <- gwr.lcr(d.a$tree~d.a$pet+d.a$rain+d.a$wqi, data = d.a, regression.points = hg, bw = lcr.bw, kernel = "bisquare", adaptive = TRUE, lambda.adjust = TRUE, cn.thresh = 30)
#summary(lcr.m$SDF$Local_CN)
#summary(lcr.m$SDF)
# extract the SPDF for analysis
m.gr3 = SpatialPointsDataFrame(lcr.m$SDF,data.frame(lcr.m$SDF), proj4string = os.proj)
```

Having tested for and identified local collinearity, a LCR-GWR was specified. This applies a GW regression but with a locally-compensated ridge term and fits local ridge regressions with their own ridge parameters (i.e., the ridge parameter varies across space), but only does this at locations where the local CN is above a user-specified threshold. In this case the CN threshold was specified as 30. An optimal adaptive bandwidth of `r lcr.bw` data points was again determined using a cross-validation procedure. Figures 6 and 7 show the spatial distribution of the original GWR coefficients, those determined using a LCR-GWR and a map of the differences between the two, for PET and for Wilderness Quality Index. In both cases there are large and potentially important differences between the coefficient estimates from the GWR and those from the LCR-GWR.

```{r, fig.width=9, echo = F, eval=T, cache = T,  message=F, warning=F, fig.cap=fig$cap("", "The coefficient estimates of the degree to which mean annual PET predicts *Acer campestre* arising from the original GWR, a locally compensated ridge GWR and a map of GWR minus LCR-GWR coefficients.")}

# Map Original GWR coefficient estimates and LCR GWR ones
# PET and GW coefs for PET 
par(mar = c(1,1,1,1))
par(mfrow = c(1,3)) 
cx <- 0.4
sh = auto.shading(m.gr2$d.a.pet, n = 7, 
	cols = brewer.pal(7, "Blues"))
plot(os10m, lwd = 0.5, add = F)
choropleth(m.gr2, m.gr2$d.a.pet, sh, pch = 19, cex = cx, add = T)
choro.legend(px = "topright", sh = sh, box.col = NA, fmt = "%2.2f", cex = 0.7, title = "GWR Coeff")
tit <- sprintf("GWR PET Coeffs")
title(tit)

plot(os10m, lwd = 0.5, add = F)
choropleth(m.gr3, m.gr3$d.a.pet, sh, pch = 19, cex = cx, add = T)
choro.legend(px = "topright", sh = sh, box.col = NA, fmt = "%2.2f", cex = 0.7, title = "LCR-GWR Coeff")
tit <- sprintf("LCR GWR PET Coeffs")
title(tit)

val <- (m.gr2$d.a.pet - m.gr3$d.a.pet)
sh = auto.shading(val, n = 7, 
	cols = brewer.pal(7, "Spectral"))
plot(os10m, lwd = 0.5, add = F)
choropleth(m.gr2, val, sh, pch = 19, cex = cx, add = T)
choro.legend(px = "topright", sh = sh, box.col = NA, fmt = "%2.2f", cex = 0.7, title = "Difference")
tit <- sprintf("LCR GWR vs GWR PET Coeffs")
title(tit)
```

```{r, fig.width=9, echo = F, eval=T, cache = T,  message=F, warning=F, fig.cap=fig$cap("", "The coefficient estimates of the degree to which Wilderness Quality Index predicts *Acer campestre* arising from the original GWR, a locally compensated ridge GWR and a map of GWR minus LCR-GWR coefficients.")}

# WQI and GW coefs for WQI
par(mar = c(1,1,1,1))
par(mfrow = c(1,3)) 
cx <- 0.4
sh = auto.shading(m.gr2$d.a.wqi, n = 7, 
	cols = brewer.pal(7, "Greens"))
plot(os10m, lwd = 0.5, add = F)
choropleth(m.gr2, m.gr2$d.a.wqi, sh, pch = 19, cex = cx, add = T)
txt=expression(paste("GW Coeff: ",B[PET]))
choro.legend(px = "topright", sh = sh, box.col = NA, fmt = "%2.2f", cex = 0.7, title = txt)
tit <- sprintf("GWR WQI Coeffs")
title(tit)

plot(os10m, lwd = 0.5, add = F)
choropleth(m.gr3, m.gr3$d.a.wqi, sh, pch = 19, cex = cx, add = T)
txt=expression(paste("LCR GW Coeff: ",B[PET]))
choro.legend(px = "topright", sh = sh, box.col = NA, fmt = "%2.2f", cex = 0.7, title = txt)
tit <- sprintf("LCR GWR WQI Coeffs")
title(tit)

val <- (m.gr2$d.a.wqi - m.gr3$d.a.wqi)
sh = auto.shading(val, n = 7, 
	cols = brewer.pal(7, "Spectral"))
plot(os10m, lwd = 0.5, add = F)
choropleth(m.gr2, val, sh, pch = 19, cex = cx, add = T)
txt= "Diff"
choro.legend(px = "topright", sh = sh, box.col = NA, fmt = "%2.2f", cex = 0.7, title = txt)
tit <- sprintf("LCR GWR vs GWR WQI Coeffs")
title(tit)
```

# 5. Discussion

In this paper a series of analyses were undertaken to demonstrate the application and value of explicitly spatial analyses, focusing on GWR, and the need to reconsider common assumptions in a-spatial data analyses. Local statistical models were developed in order to test for spatial non-stationarity, in contrast to standard, a-spatial, statistical approaches that assume the relationships between factors to be the same everywhere. Additionally, the paper highlights the importance of considering and testing for local collinearity especially in spatial non-stationarity models such as GWR, even where none is found to exist globally. In this analysis very strong evidence for local collinearity was found when the data were tested using local collinearity diagnostics. In all applications of GWR published in the literature, this critical step has been missed. Where local collinearity is found, locally-compensated models such as LCR-GWR can be applied and Brunsdon et al. (2012) propose a variety of these. A LCR-GWR only fits local ridge regressions at locations where the local Condition Number is above a user-specified threshold. Gollini et al (2015) review LCR-GWR with respect to alternative approaches for handling collinearity, but whatever adapted GWR model is applied, they all have the potential provide more accurate local coefficient estimates in the presence of collinearity than that found with a standard GWR model (Brunsdon et al. 2012).

The Geographically Weighted (GW) paradigm offers an attractive and coherent framework for many areas of applied geographical analysis. Geographically Weighted approaches, testing for spatial non-stationarity, are in contrast to standard, a-spatial, statistical approaches that assume the relationships between factors to be the same everywhere. They reflect Tobler’s 1st Law of Geography and an understanding of the world when it is viewed through ‘Geography Goggles’. These promote a vision in which the wearer is interested in how and where things vary, does not expect (statistical) relationships to be same everywhere, does not consider the world to be normally distributed especially in space, but rather expects processes, relationships, processes, trends etc. to vary spatially and to find clusters, hotspots, coldspots, etc.

These ideas are not new: quantitative geography in 1980s identified the need to move away from the whole map statistics, particularly Stan Openshaw’s group at Newcastle, UK and Julian Besag’s at Durham, UK but also Luc Anselin at Arizona, USA. Similar ideas in arose in ecology and the advent of spatial ecology. But it is important to re-state them now for a number of reasons. First, all data are spatial now (well perhaps not quite all!) but with advent of ubiquitous GPS, most records, datasets and data points have location attached to them. Second, location is not just another variable precisely because of spatial heterogeneity and spatial dependence observed in many processes, with the result that many phenomena are *not* constant or randomly distributed, as predicated by classic statistical models. Third, the need to think spatially and to and to consider the spatial dimensions in a different way is given further salience by the increased access to and use of very powerful GIS software. This is increasingly resulting in instances of poor and inappropriate use of very powerful tools, but that is another story (see Comber et al., 2015). Finally, we simply observe that geography goggles are not usually worn by researchers working in many areas of applied ecology and bio-geography, especially in conservation, environmental science and remote sensing, where the whole map statistic persists.

# 6. Conclusions

Nearly all data are spatial nowadays with most being collected some*where*. However, locational attributes of data require careful consideration as many of the landscape processes (both environmental and social) that we observe, measure and model do not exhibit spatial constancy (stationarity) or randomness. The 
spatial heterogeneity and spatial dependence of landscape processes is therefore problematic for classic statistical models and inference when nearby features are found to be correlated, and local, spatially explicit models have been proposed by many authors to handle spatial-non-stationarity. However, local models such as the moving window or kernel used by GWR may be subject to *local* collinearity amongst variables, even when none is observed *globally*. This paper demonstrates the use of a locally compensated ridge GWR to overcome local collinearity where found. It demonstrates how carefully considered spatially explicit statistical models can be used examine the spatial variation in processes and relationships and to account for spatial autocorrelation.

# Acknowledgements

The authors would like to thank Mark O’Connell and the 4th Spatial Ecology and Conservation conference for the opportunity to develop these ideas. This research was also supported by the Biotechnology and Biological Sciences Research Council (BBSRC BB/J004308/1) for the North Wyke Farm Platform.

# References

Akaike H (1973). Information Theory and an Extension of the Maximum Likelihood Principle. In BN Petrov, F Csaki (eds.), *2nd Symposium on Information Theory*, pp. 267–281. Akademiai Kiado, Budapest.

Anselin L (1988). *Spatial Econometrics: Methods and Models*, Dordrecht: Kluwer Academic 
Assunção R, 2003, Space varying coefficient models for small area data, *Environmetrics*, 14: 453-473

Berry BJL, Marble DF (1968) *Spatial Analysis*, Englewood Cliffs, NJ: Prentice Hall

Belsley DA, Kuh E, Welsch RE (1980). *Regression Diagnostics: Identifying Influential Data and Sources of Collinearity*. John Wiley & Sons.

Besag, J (1974). Spatial interaction and the statistical analysis of lattice systems (with discussion). *Journal of the Royal Statistical Society B* 36: 192-225.

Bowman A (1984). An Alternative Method of Cross-Validation for the Smoothing of Density Estimates. *Biometrika*, 71, 353–360.

Brunsdon C, Charlton M, Harris P (2012). Living with Collinearity in Local Regression Models. In *Proceedings of the 10th International Symposium on Spatial Accuracy Assessment in Natural Resources and Environmental Sciences*. Brasil.

Brunsdon C, Fotheringham AS, Charlton M (1996). Geographically Weighted Regression: A Method for Exploring Spatial Nonstationarity. *Geographical Analysis*, 28, 281–289.

Carver S, Comber A, McMorran R and Nutter S. (2012). A GIS model for mapping spatial patterns and distribution of wild land in Scotland. *Landscape and Urban Planning*, 104 (2012) 395–409.

Casetti E (1982). Drift analysis of regression parameters: An application to the investigation of fertility development relations, *Modeling & Simulation* 13: 961-966.

Chiles, J.-P. and P. Delfiner (1999). *Geostatistics - modelling spatial uncertainty.*

Chorley R, Haggett P (1967). *Models in Geography*, London: Methuen

Comber AJ, Brunsdon C, Green E. (2008). Using a GIS-based network analysis to determine urban greenspace accessibility for different ethnic and religious groups. *Landscape and Urban Planning*, 86: 103–114.

Comber AJ, Carver S, Fritz S, McMorran R, Washtell J and Fisher P. (2010). Different methods, different wilds: evaluating alternative mappings of wildness using Fuzzy MCE and Dempster Shafer MCE. *Computers, Environment and Urban Systems*, 34: 142-152

Comber A, Dickie J, Jarvis C, Phillips M and Tansey K, (2015). Locating bioenergy facilities using a modified GIS-based location-allocation-algorithm: considering the spatial distribution of resource supply. *Applied Energy*, 154: 309-316.

Coudun, C., Gégout, J.-C., Piedallu, C. and Rameau, J.-C. (2006), Soil nutritional factors improve models of plant species distribution: an illustration with Acer campestre (L.) in France. Journal of Biogeography, 33: 1750–1763

Cressie, N (1993). *Statistics for Spatial Data*. New Jersey, John Wiley.

Cressie, N and Wilke CK (2011) *Statistics for Spatio-Temporal Data*, New Jersey, John Wiley.

Dormann, C. F., Elith, J., Bacher, S., Buchmann, C., Carl, G., Carré, G., Marquéz, J. R. G., Gruber, B., Lafourcade, B., Leitão, P. J., Münkemüller, T., McClean, C., Osborne, P. E., Reineking, B., Schröder, B., Skidmore, A. K., Zurell, D. and Lautenbach, S. (2013). Collinearity: a review of methods to deal with it and a simulation study evaluating their performance. *Ecography* 36: 27-46.

Dormann, C. F., Schymanski, S. J., Cabral, J., Chuine, I., Graham, C., Hartig, F., Kearney, M., Morin, X., Römermann, C., Schröder, B. and Singer, A. (2012), Correlation and process in species distribution models: bridging a dichotomy. *Journal of Biogeography*, 39: 2119–2131

Dormann, C. F. (2007). Effects of incorporating spatial autocorrelation into the analysis of species distribution data. *Global Ecology and Biogeography*, 16(2), 129-138.

Fisher (1935). Statistical Tests, *Nature* 136, 474.

Foody, G.M. (2004) Spatial non-stationarity and scale dependency in the relationship between species richness and environmental determinants for the sub-Saharan endemic avifauna. *Global Ecology and Biogeography*, 13, 315–320.

Fotheringham, A. S., Brunsdon, C., & Charlton, M. (2002). Geographically weighted regression: the analysis of spatially varying relationships. John Wiley & Sons.

Fujita, M. (1989). Urban Economic Theory, Land Use and City Size, Cambridge: Cambridge University Press.

Gandin, L. S. (1965). *Objective analysis of meteorological fields*.Israel Program for Scientific Translation, Jerusalem.

Gelfand AE, Kim HJ, Sirmans CJ, Banerjee S (2003). Spatial modelling with spatially varying coefficient processes. *Journal of American Statistical Association* 98: 387-396

Gollini, I., Lu, B., Charlton, M., Brunsdon, C., & Harris, P. (2015). GWmodel: an R package for exploring spatial heterogeneity using geographically weighted models. *Journal of Statistical Software*, 63 (17), 1–50.

Goovaerts P (2009). Medical Geography: A promising field of application for geostatistics. *Mathematical Geosciences* 41: 243-264

Gorr WL, Olligschlaeger AM (1994). Weighted Spatial Adaptive Filtering: Monte Carlo Studies and Application to Illicit Drug Market Modeling. *Geographical Analysis*, 26: 67-87.

Guisan A, Tingley R, Baumgartner JB, Naujokaitis-Lewis I, Sutcliffe PR, Tulloch AIT, Regan TJ, Brotons L, McDonald-Madden E, Mantyka-Pringle C, Martin TG, Rhodes JR, Maggini R, Setterfield SA, Elith J, Schwartz MW, Wintle BA, Broennimann O, Austin M, Ferrier S, Kearney MR, Possingham HP, Buckley YM (2013). Predicting species distributions for conservation decisions. *Ecology Letters*, 16: 1424 - 1435

Harris P, Brunsdon C, Fotheringham AS (2011). Links, comparisons and extensions of the geographically weighted regression model when used as a spatial predictor. *Stochastic Environmental Research and Risk Assessment* 25: 123-138

Hijmans, Robert J., and Jane Elith. *Species distribution modeling with R*. (2016). http://www.idg.pl/mirrors/CRAN/web/packages/dismo/vignettes/sdm.pdf

Hoerl AE (1962). Application of Ridge Analysis to Regression Problems. *Chemical Engineering Progress*, 58(3), 54–59.

Hoerl AE, Kennard RW (1970). Ridge Regression: Biased Estimation for Nonorthogonal Problems. *Technometrics*, 12(1), 55–67.

Honrado, J.P., Pereira, H.M., Guisan, A. (2016), Fostering integration between biodiversity monitoring and modelling, *Journal of Applied Ecology*, 53: 1299-1304.

Journel, A. G. and C. J. Huijbregts (1978). *Mining geostatistics*. London, Academic Press.

Keith SA, Kerswell AP, and Connolly SR (2014). Global diversity of marine macroalgae: environmental conditions explain less variation in the tropics. *Global ecology and Biogeography*, 23(5), 517-529.

Kery M., B. Gardner, and C. Monnerat (2010). Predicting species distributions from checklist data using site-occupancy models. *Journal of Biogeography*. 37: 1851–1862

Kolmogorov, AN (1941). Interpolirovanie i ekstrapolirovanie statsionarnykh sluchainykh posledovatel' nostei (Interpolated and extrapolated stationary random sequences). Izvestya Akademiya Nauk SSSR, *Seria Matematicheskaya* 5(1).

Krige, DG (1951). A statistical approach to some basic mine valuation problems on the Witwatersrand. *Journal of the Chemical, Metallurgical and Mining Society of South Africa* 52: 119-139.

Kuiters, A. T., van Eupen, M., Carver, S., Fisher, M., Kun, Z., & Vancura, V. (2013). *Wilderness register and indicator for Europe. Final Report*, http://ec.europa.eu/environment/nature/natura2000/wilderness/pdf/Wilderness_register_indicator.pdf

Lu, B., Harris, P., Charlton, M., & Brunsdon, C. (2014). The GWmodel R package: further topics for exploring spatial heterogeneity using geographically weighted models. Geo-spatial Information Science, 17(2), 85-101.

Matérn (1960). Spatial variation: stochastic models and their applications to problems in forest surveys and other sampling investigations. *Meddelanden fran Statens Skogforskningsinstitut*, 49: 1-144.

Matheron, G. (1963). Principles of geostatistics. *Economic Geology* 58: 1246-1266.

Meloun, M. et al. 2002. Crucial problems in regression modelling and their solutions. *Analyst* 127: 433–450.

O’Brien RM (2007). A Caution Regarding Rules of Thumb for Variance Inflation Factors. *Quality & Quantity*, 41(5), 673–690.

Openshaw S and Taylor PJ (1979). A million or so correlation coefficients: three experiments on the modifiable areal unit problem. In: *Statistical methods in the spatial sciences*, (ed) N. Wrigley, Pion: London, 127-144

Pearman P.B., Guisan A., Broennimann O., Randin C.F., (2008). Niche dynamics in space and time. *Trends in Ecology and Evolution*, 23: 149-158.

Phillips S.J. and J. Elith, 2011. Logistic methods for resource selection functions and presence-only species distribution models, AAAI (Association for the Advancement of Artificial Intelligence), San Francisco, USA.

Phillips, S.J., M. Dudik, J. Elith, C.H. Graham, A. Lehmann, J. Leathwick, and S. Ferrier. 2009. Sample selection bias and presence-only distri- bution models: implications for background and pseudo-absence data. Ecological Applications 19: 181-197.

Rocchini, D., Andreo, V., Förster, M., Garzon-Lopez, C.X., Gutierrez, A.P, Gillespie, T.W., Hauffe, H.C., He, K.S., Kleinschmit, B., Mairota, P., Marcantonio, M., Metz, M., Nagendra, H., Pareeth, S., Ponti, L., Ricotta, C., Rizzoli, A., Schaab, G., Zebisch, M., Zorer, R., Neteler, M. (2015). Potential of remote sensing to predict species invasions: A modelling perspective. *Progress in Physical Geography*, 39: 283-309.

Rocchini, D., Petras, v., Petrasova, A,, Chemin, Y., Ricotta, C., Frigeri, A., Landa, M., Marcantonio, M., Bastin, L., Metz, M., Delucchi, L., Neteler, M. (2016 in press). Spatio-ecological complexity measures in GRASS GIS. *Computers & Geosciences*.

Roll U, Geffen E, and Yom-Tov Y (2015). Linking vertebrate species richness to tree canopy height on a global scale. *Global Ecology and Biogeography*, 24(7), 814-825.

Tanguy, M.; Dixon, H.; Prosdocimi, I.; Morris, D. G.; Keller, V. D. J. (2015). *Gridded estimates of daily and monthly areal rainfall for the United Kingdom (1890-2014) [CEH-GEAR]*. NERC Environmental Information Data Centre. http://doi.org/10.5285/f2856ee8-da6e-4b67-bedb-590520c77b3c

Tobler WR (1970). A computer movie simulating urban growth in the Detroit region. *Economic Geography*, 46: 234-240

Trabucco A, and Zomer RJ (2009). *Global Aridity Index (Global-Aridity) and Global Potential Evapo-Transpiration (Global-PET) Geospatial Database. CGIAR Consortium for Spatial Information*. Published online, available from the CGIAR-CSI GeoPortal at: http://www.csi.cgiar.org

VanDerWal J., L.P. Shoo, C. Graham and S.E. Williams, 2009. Selecting pseudo-absence data for presence-only distribution modeling: how far should you stray from what you know? *Ecological Modelling*, 220: 589-594.

Ward G., T. Hastie, S.C. Barry, J. Elith and J.R. Leathwick, 2009. Presence- 
only data and the EM algorithm. Biometrics 65: 554-563.

Wall MM (2004). A close look at the spatial structure implied by the CAR and SAR models, *Journal of Statistical Planning and Inference*, 121: 311-24

Wang Q, Ni J, and Tenhunen J (2005). Application of a geographically-weighted regression analysis to estimate net primary production of Chinese forest ecosystems. *Global Ecology and Biogeography*, 14(4), 379-393.

Wheeler D (2007). Diagnostic Tools and a Remedial Method for Collinearity in Geographically Weighted Regression. *Environment and Planning A*, 39(10), 2464–2481.

Wheeler D (2009). Simultaneous Coefficient Penalization and Model Selection in Geographically Weighted Regression: the Geographically Weighted Lasso. *Environment and Planning A*, 41(3), 722–742.

Wheeler D (2013). Geographically Weighted Regression. In M Fischer, P Nijkamp (eds.), *Handbook of Regional Science*. Springer-Verlag.

Wheeler D, Tiefelsdorf M (2005). Multicollinearity and Correlation among Regression Co-efficients in Geographically Weighted Regression. *Journal of Geographical Systems*, 7(2), 161–187.

Yaglom AM (1955). Correlation theory of processes with random stationary nth increments. *Matematicheskii Sbornik*, 37: 141-196.

Zou, H., & Hastie, T. (2005). Regularization and variable selection via the elastic net. *Journal of the Royal Statistical Society: Series B (Statistical Methodology)*, 67(2), 301-320.